Preparing your manuscript

### 2025-02-17 article/article.qmd
weighted mean

We estimate storm tracks by calculating transient eddy kinetic energy (EKE) using a Fast Fourier Transform high-pass filter applied to daily horizontal wind at 300 hPa, with a 6-day cutoff for each model.

@@

```{r}
#| label: fig-serie
#| fig-width: 8
#| fig-height: 5
#| fig.cap: "Mean frequency of $\\Delta T$ for the historical period (1850-2014) and SSP5-8.5 (2015-2100) over land and sea for different latitude bands. Left to each main figure the intermodel distribution is shown for the period 2080-2100."

@@

  _[, let(land = MaskLand(lon, lat),
          region = fcase(lat %between% c(25, 65), "NH",
                         lat %between% c(-65, -25), "SH",
                         default = NA))] |> 
  _[, let(land = factor(land, labels = c("Sea", "Land")))] |> 
  na.omit() |> 
  _[, .(mean_eke = mean(eke),
        mean_deltat = mean(deltap97.5)), by = .(region, land, model)] |> 
  ggplot(aes(mean_eke, mean_deltat)) +
  geom_point(aes(color = region, shape = land), size = 2) +
  geom_smooth(aes(colour = region), method = "lm", se = FALSE) +
  scale_color_manual(values = c("cyan4", "orange")) +
  labs(x = latex2exp::TeX("EKE change ($m^2 s^{-2}$)"), y = latex2exp::TeX("$\\Delta T$ frequency change (days/year)"),
       color = NULL, shape = NULL) +
  theme_minimal() +
  theme(legend.position = "inside",
        legend.position.inside = c(0.83, 0.08),
        legend.direction = "horizontal",
        legend.background = element_rect(fill = "#fbfbfb", color = "#fbfbfb"),
        legend.margin = margin(-3, 0, 0, 0, unit = "mm"),
        panel.background = element_rect(fill = "#fbfbfb", color = NA))

deltat_mean <- data |> 
  _[(model %in% models_in_gadi) & is.finite(change)] |>
  dcast(lat + lon + model ~ var, value.var = "diff") |> 
  _[, let(land = MaskLand(lon, lat))] |> 
  _[, let(land = factor(land, labels = c("Sea", "Land")))] |> 
  _[, .(mean_dt = mean(deltap97.5-1, na.rm = TRUE)), by = .(lat, land)]

mean_trend_dt <- data |> 
  _[(model %in% models_in_gadi) & is.finite(change)] |>
  dcast(lat + lon + model ~ var, value.var = "diff") |> 
  _[, let(land = MaskLand(lon, lat))] |> 
  _[, let(land = factor(land, labels = c("Sea", "Land")))] |> 
  _[, .(mean_eke = mean(eke, na.rm = TRUE)), by = .(lat, land, model)] |> 
  ggplot(aes(mean_eke-1, lat)) +
  geom_hline(yintercept = 0, color = "grey80") +
  geom_vline(xintercept = 0, color = "grey80") +
  geom_path(aes(group = model), color = "grey70", alpha = 0.7) +
  stat_summary(orientation = "y", geom = "line", aes(color = "EKE")) +
  geom_path(data = deltat_mean, aes(x = mean_dt, color = "DeltaT")) +
  scale_color_manual(values = c("DeltaT" = "red4", "EKE" = "black"),
                     labels = c(latex2exp::TeX("$\\Delta T$"), "EKE")) +
  geom_rect(aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "#fbfbfb",
            data = annotation,
            inherit.aes = FALSE) +
  annotate(geom = "rect", ymin = c(-90, -20, 90), ymax = c(-65, 20, 65), xmin = -Inf, xmax = Inf,
           fill = "grey93", alpha = 0.7) +
  scale_y_latitude(breaks = seq(-80, 80 , 20)) +
  facet_wrap(~land) +
  tagger::tag_facets(tag_pool = c("b", "c")) +
  labs(x = "Change", color = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom",
        panel.background = element_rect(fill = "#fbfbfb", color = NA),
        strip.text.x = element_text(size = 12),
        plot.margin = unit(c(0, 0 , 0, -0.40), units = "cm"),
        tagger.panel.tag.text = element_text(size = 12))

>>>>

To estimate the storm track location we use the transient eddy kinetic energy (EKE), used by many others to estimate storm tracks [@ogorman2010; @chemke2022]. We calculated EKE using a Fast Fourier Transform high-pass filter applied to daily horizontal wind at 300 hPa, with a 6-day cutoff for each model.

@@

```{r}
#| label: fig-serie
#| fig-width: 8
#| fig-height: 5
#| fig.cap: "Multimodel mean frequency of $\\Delta T$ for the historical period (1850-2014) and SSP5-8.5 (2015-2100) over land and sea for different latitude bands. Left to each main figure the intermodel distribution is shown for the period 2080-2100."

@@

  _[, let(land = MaskLand(lon, lat),
          region = fcase(lat %between% c(25, 65), "NH",
                         lat %between% c(-65, -25), "SH",
                         default = NA))] |> 
  _[, let(land = factor(land, labels = c("Sea", "Land")))] |> 
  na.omit() |> 
  _[, .(mean_eke = weighted.mean(eke, w = cos(lat*pi/180)),
        mean_deltat = weighted.mean(deltap97.5, w = cos(lat*pi/180))), by = .(region, land, model)] |> 
  ggplot(aes(mean_eke, mean_deltat)) +
  geom_point(aes(color = region, shape = land), size = 2) +
  geom_smooth(aes(colour = region), method = "lm", se = FALSE) +
  scale_color_manual(values = c("cyan4", "orange")) +
  labs(x = latex2exp::TeX("EKE change ($m^2 s^{-2}$)"), y = latex2exp::TeX("$\\Delta T$ frequency change (days/year)"),
       color = NULL, shape = NULL) +
  theme_minimal() +
  theme(legend.position = "inside",
        legend.position.inside = c(0.83, 0.08),
        legend.direction = "horizontal",
        legend.background = element_rect(fill = "#fbfbfb", color = "#fbfbfb"),
        legend.margin = margin(-3, 0, 0, 0, unit = "mm"),
        panel.background = element_rect(fill = "#fbfbfb", color = NA))

deltat_mean <- data |> 
  _[(model %in% models_in_gadi) & is.finite(change)] |>
  dcast(lat + lon + model ~ var, value.var = "diff") |> 
  _[, let(land = MaskLand(lon, lat))] |> 
  _[, let(land = factor(land, labels = c("Sea", "Land")))] |> 
  _[, .(mean_dt = mean(deltap97.5, na.rm = TRUE)), by = .(lat, land)]

mean_trend_dt <- data |> 
  _[(model %in% models_in_gadi) & is.finite(change)] |>
  dcast(lat + lon + model ~ var, value.var = "diff") |> 
  _[, let(land = MaskLand(lon, lat))] |> 
  _[, let(land = factor(land, labels = c("Sea", "Land")))] |> 
  _[, .(mean_eke = mean(eke, na.rm = TRUE)), by = .(lat, land, model)] |> 
  ggplot(aes(mean_eke, lat)) +
  geom_hline(yintercept = 0, color = "grey80") +
  geom_vline(xintercept = 0, color = "grey80") +
  geom_path(aes(group = model), color = "grey70", alpha = 0.7) +
  stat_summary(orientation = "y", geom = "line", aes(color = "EKE")) +
  geom_path(data = deltat_mean, aes(x = mean_dt, color = "DeltaT")) +
  scale_color_manual(values = c("DeltaT" = "red4", "EKE" = "black"),
                     labels = c(latex2exp::TeX("$\\Delta T$"), "EKE")) +
  geom_rect(aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "#fbfbfb",
            data = annotation,
            inherit.aes = FALSE) +
  annotate(geom = "rect", ymin = c(-90, -20, 90), ymax = c(-65, 20, 65), xmin = -Inf, xmax = Inf,
           fill = "grey93", alpha = 0.7) +
  scale_y_latitude(breaks = seq(-80, 80 , 20)) +
  facet_wrap(~land) +
  tagger::tag_facets(tag_pool = c("b", "c")) +
  labs(x = "Change", color = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom",
        panel.background = element_rect(fill = "#fbfbfb", color = NA),
        strip.text.x = element_text(size = 12),
        plot.margin = unit(c(0, 0 , 0, -0.40), units = "cm"),
        tagger.panel.tag.text = element_text(size = 12))
